---
title: "A Timed Test"
output: html_notebook
Author: Igor Pupko
test start: 19:46
---

There were a total of 100 participants who gave blood to screen for specific metabolites in blood, that were previously unknown. Additionally, they filled questionnaires where they recorded their biological sex, age and BMI. It is a task of researcher to identify the characteristics of those new molecules. 


#Researcher Note. Enable required packages to prepare workspace

```{r}

library(ggplot2)#Package for plotting and graphics
library(simstudy)#Handy package for simulating data sets
library(dplyr)
library(tidyverse)
```

Workspace was prepared. 

#simmulating a dataset. Creating a variables that will be dependent on participant sex, age and BMI.
```{r}
#Create a dummy data set
#Creating dummy age variable
def <- defData(varname = "age", dist = "normal", formula = 24,
               variance = 2)
#Creating dummy Sex variable, using age as seed
def <- defData(def, varname = "Sex", dist = "binary",
               formula = "-2 + age * 0.083", link = "logit")


```
