---
title: "A Timed Test"
output:
  html_document:
    df_print: paged
Author: Igor Pupko
test start: '19:46'
---

There were a total of 100 participants who gave blood to screen for specific metabolites in blood, that were previously unknown. Additionally, they filled questionnaires where they recorded their biological sex, age and BMI. It is a task of researcher to identify the characteristics of those new molecules. 


#Researcher Note. Enable required packages to prepare workspace

```{r}

library(ggplot2)#Package for plotting and graphics
library(simstudy)#Handy package for simulating data sets
library(dplyr)
library(tidyverse)
```

Workspace was prepared. 

#simulating a dataset. Creating a variables that will be dependent on study participant's sex, age and BMI.
```{r}
#Create a dummy data set
#Creating dummy age variable
def <- defData(varname = "age", dist = "normal", formula = 24,
               variance = 2)
#Creating dummy Sex variable, using age as seed
def <- defData(def, varname = "Sex", dist = "binary",
               formula = "-2 + age * 0.083", link = "logit")
#Creating Dummy BMI Variable
def <- defData(def, varname = "BMI", dist = "normal",
               formula = 22, variance=2)
#Creating Dummy Metabolic Variable 1, depends on BMI and age
def <- defData(def, varname = "Metabolite_1", dist = "normal",
                formula = "1.5 - 0.3 * age + 2 * BMI", link = "log")
#Creating Dummy Metabolic Variable 2, depends on BMI and sex
def <- defData(def, varname = "Metabolite_2", dist = "normal",
                formula = "4.8 - 0.1 * Sex + 21.7 * BMI", link = "log")
#Creating Dummy Metabolic Variable 3, depends on BMI and sex
  def <- defData(def, varname = "Metabolite_3", dist = "normal",
                formula = "3.1 + 15.1 * BMI + 2.4 * Sex", variance = 15)
#create a data table
set.seed(42) #Random number as seed,a pop-culture reference
dummy<-genData(100,def) #Generate dummy data set
dummy$age<-round(dummy$age) # For simpler data

```
#Simulted date table with participant information was created.
Creating a local txt backup.
```{r}
#Create a backup text file for the table
write.table(dummy,file="dummy.txt", quote = FALSE)
```
Backup Created.

Begin description of dataset.
```{r}

```

