---
title: "A Timed Test"
output:
  html_document:
    df_print: paged
Author: Igor Pupko
test start: '19:46'
---

There were a total of 100 participants who gave blood to screen for specific metabolites in blood, that were previously unknown. Additionally, they filled questionnaires where they recorded their biological sex, age and BMI. It is a task of researcher to identify the characteristics of those new molecules. 


#Researcher Note. Enable required packages to prepare workspace

```{r}

library(ggplot2)#Package for plotting and graphics
library(simstudy)#Handy package for simulating data sets
library(dplyr)
library(tidyverse)
```

Workspace was prepared. 

#simulating a dataset. Creating a variables that will be dependent on study participant's sex, age and BMI.
```{r}
#Create a dummy data set
#Creating dummy age variable
def <- defData(varname = "age", dist = "normal", formula = 24,
               variance = 2)
#Creating dummy Sex variable, using age as seed
def <- defData(def, varname = "Sex", dist = "binary",
               formula = "-2 + age * 0.083", link = "logit")
#Creating Dummy BMI Variable
def <- defData(def, varname = "BMI", dist = "normal",
               formula = 22, variance=2)
#Creating Dummy Metabolic Variable 1, depends on BMI and age
def <- defData(def, varname = "Metabolite_1", dist = "normal",
                formula = "1.5 - 0.3 * age + 2 * BMI", link = "log")
#Creating Dummy Metabolic Variable 2, depends on BMI and sex
def <- defData(def, varname = "Metabolite_2", dist = "normal",
                formula = "4.8 - 0.1 * Sex + 21.7 * BMI", link = "log")
#Creating Dummy Metabolic Variable 3, depends on BMI and sex
  def <- defData(def, varname = "Metabolite_3", dist = "normal",
                formula = "3.1 + 15.1 * BMI + 2.4 * Sex", variance = 15)
#create a data table
set.seed(42) #Random number as seed,a pop-culture reference
dummy<-genData(100,def) #Generate dummy data set
dummy$age<-round(dummy$age) # For simpler data

```
#Simulted date table with participant information was created.
Creating a local txt backup.
```{r}
#Create a backup text file for the table
write.table(dummy,file="dummy.txt", quote = FALSE)
```
Backup Created.

Begin description of data table. Write results to txt file.
```{r}
original_dataset<-dummy
variables<-colnames(original_dataset)
original_dataset$id<-NULL #remove ids to prevent interference with data
dfname <- deparse(substitute(dummy)) #variable with the name of data table
#make sex variables categorical
original_dataset$Sex[original_dataset$Sex==1]<-"male"
original_dataset$Sex[original_dataset$Sex==0]<-"female"
#Create a log text file with the description of data table
fileConn=paste0("README_", dfname,"_decription.txt") #Create a log file
cat("Data that is processed is:", dfname,"\n",file=fileConn, sep= " ",  append = TRUE) #Identification of source data
cat("\n Number of individuals:", dim(original_dataset)[1], "\n",file=fileConn, sep= " ",  append = TRUE) #Number of individuals
cat("\n Number of Variables:", dim(original_dataset)[2], "\n",file=fileConn, sep= " ",  append = TRUE) #Number of variables
#Calculate Number of Males
cat("\n Number of Males:", length(original_dataset$Sex[original_dataset$Sex=="male"]), "\n",file=fileConn, sep= " ",  append = TRUE)
#Calculate Number of Females
cat("\n Number of Females:", length(original_dataset$Sex[original_dataset$Sex=="female"]), "\n",file=fileConn, sep= " ",  append = TRUE)


for (i in colnames(original_dataset)) {
  cat("\n Mean of ",i,":", mean(original_dataset[[i]]), "\n",file=fileConn, sep= " ",  append = TRUE)
  cat("\n Min of ",i,":", min(original_dataset[[i]]), "\n",file=fileConn, sep= " ",  append = TRUE)
  cat("\n Max of ",i,":", max(original_dataset[[i]]), "\n",file=fileConn, sep= " ",  append = TRUE)
  cat("\n Range of ",i,":", range(original_dataset[[i]]), "\n",file=fileConn, sep= " ",  append = TRUE)
}




#BMI Distribution Plot
pdf((paste0("Distribution plot of BMI.pdf")),width = 8,  height = 11)
  # The Distribution Plot
ggplot(original_dataset, aes(x=BMI,fill=Sex)) + geom_density(alpha=.5)
#ggplot(dat, aes(x=rating, fill=cond)) + geom_density(alpha=.3)
dev.off()

#age Distribution Plot
pdf((paste0("Distribution plot of age.pdf")),width = 8,  height = 11)
  # The Distribution Plot
ggplot(original_dataset, aes(x=age,fill=Sex)) + geom_density(alpha=.5)
#ggplot(dat, aes(x=rating, fill=cond)) + geom_density(alpha=.3)
dev.off()



#Metabolite 1 Distribution Plot
pdf((paste0("Distribution plot of Metabolite 1.pdf")),width = 8,  height = 11)
  # The Distribution Plot
ggplot(original_dataset, aes(x=Metabolite_1,fill=Sex)) + geom_density(alpha=.5)
#ggplot(dat, aes(x=rating, fill=cond)) + geom_density(alpha=.3)
dev.off()

#Metabolite 2 Plot
pdf((paste0("Distribution plot of Metabolite 2.pdf")),width = 8,  height = 11)
  # The Distribution Plot
ggplot(original_dataset, aes(x=Metabolite_2,fill=Sex)) + geom_density(alpha=.5)
#ggplot(dat, aes(x=rating, fill=cond)) + geom_density(alpha=.3)
dev.off()

#Metabolite 3 Distribution Plot
pdf((paste0("Distribution plot of Metabolite 3.pdf")),width = 8,  height = 11)
  # The Distribution Plot
ggplot(original_dataset, aes(x=Metabolite_3,fill=Sex)) + geom_density(alpha=.5)
#ggplot(dat, aes(x=rating, fill=cond)) + geom_density(alpha=.3)
dev.off()

```
Initial data description created and log file written.
Begin analysis stage.
Linear Regression and Scatter Plot.
```{r}
variables2<-original_dataset[,c(4:6)]

for (q in colnames(variables2)) { 
mx<-lm(variables2[[q]] ~ Sex + BMI + age, data = original_dataset)

sink(paste0("README_",q,"_LM_decription.txt"))
print(summary(mx))
close=fileConn
}

#Scatter Plots vs BMI
pdf((paste0("Scatter plot Metabolite 1 vs BMI.pdf")),  width = 4,  height = 4)
ggplot(dummy, aes(x=BMI, y=Metabolite_1)) + geom_point() + geom_smooth(method=lm, color="red", se=TRUE)
dev.off()

pdf((paste0("Scatter plot Metabolite 2 vs BMI.pdf")),  width = 4,  height = 4)
ggplot(dummy, aes(x=BMI, y=Metabolite_2)) + geom_point() + geom_smooth(method=lm, color="red", se=TRUE)
dev.off()

pdf((paste0("Scatter plot Metabolite 3 vs BMI.pdf")),  width = 4,  height = 4)
ggplot(dummy, aes(x=BMI, y=Metabolite_3)) + geom_point() + geom_smooth(method=lm, color="red", se=TRUE)
dev.off()

#Scatter Plots vs age

pdf((paste0("Scatter plot Metabolite 1 vs age.pdf")),  width = 4,  height = 4)
ggplot(dummy, aes(x=age, y=Metabolite_1)) + geom_point() + geom_smooth(method=lm, color="red", se=TRUE)
dev.off()

pdf((paste0("Scatter plot Metabolite 2 vs age.pdf")),  width = 4,  height = 4)
ggplot(dummy, aes(x=age, y=Metabolite_2)) + geom_point() + geom_smooth(method=lm, color="red", se=TRUE)
dev.off()

pdf((paste0("Scatter plot Metabolite 3 vs age.pdf")),  width = 4,  height = 4)
ggplot(dummy, aes(x=age, y=Metabolite_3)) + geom_point() + geom_smooth(method=lm, color="red", se=TRUE)
dev.off()

#Scatter Plots vs sex

pdf((paste0("Scatter plot Metabolite 1 vs sex.pdf")),  width = 4,  height = 4)
ggplot(dummy, aes(x=Sex, y=Metabolite_1)) + geom_point() + geom_smooth(method=lm, color="red", se=TRUE)
dev.off()

pdf((paste0("Scatter plot Metabolite 2 vs age.pdf")),  width = 4,  height = 4)
ggplot(dummy, aes(x=Sex, y=Metabolite_2)) + geom_point() + geom_smooth(method=lm, color="red", se=TRUE)
dev.off()

pdf((paste0("Scatter plot Metabolite 3 vs age.pdf")),  width = 4,  height = 4)
ggplot(dummy, aes(x=Sex, y=Metabolite_3)) + geom_point() + geom_smooth(method=lm, color="red", se=TRUE)
dev.off()
```

Analysis complete. 
From distribution plot, it seems that metabolite 3 level is higher in men.
From linear regression analysis it seems that levels of Metabolite 3 is dependent on BMI and SEX as seen through regression analysis. *As was created in simulated data.


SHAMELESS DISCLAIMER: Researcher knows a faster way to complete the analysis through use of loops, but his R packages went haywire on the moment of the beigining of the test.

Test Completion Time: 20:55 (BST)
